A Comparative Study of Game-Theoretic and Reinforcement Learning Approaches for Jam-Resilient Wireless Networks in IoT

### 2025-05-26 paper/figures/main.tex
Progress (IEEE Format)

% LaTeX template for "A Comparative Study of Game-Theoretic and Reinforcement Learning Approaches for Jam-Resilient Wireless Networks in IoT"
\documentclass[conference]{IEEEtran}
%======================== Packages ========================
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{algorithm,algpseudocode}
\usepackage{cite}
\usepackage{url}
\usepackage{booktabs}

@@

\author{
  \IEEEauthorblockN{Muhammad Umar Yaksambi}
  \IEEEauthorblockA{
    Dept. of CSE (Data Sciences)\\
    RV College Of Engineering\\
    Bengaluru, India\\
    Email: umaryaksambi@gmail.com
  }
  \and
  \IEEEauthorblockN{Anuj Devapura}
  \IEEEauthorblockA{
    Dept. of CSE (Data Sciences)\\
    RV College Of Engineering\\
    Bengaluru, India\\
    Email: anujdevapura@gmail.com
  }
  \and
  \IEEEauthorblockN{Sadashiv Todakar}
  \IEEEauthorblockA{
    Dept. of CSE (Data Sciences)\\
    RV College Of Engineering\\
    Bengaluru, India\\
    Email: sadashivtodakar@gmail.com
  }
}

@@

%--- Figures ---
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.45\textwidth]{interval_avg_attacker_payoff_filt_topologyRandom_Erdos_Renyinum_nodes10connectivity_param0p4_matchup_matrix_heatmap.png}
  \caption{Interval-averaged attacker payoff across strategy matchups.}
  \label{fig:atk_payoff_heatmap}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.45\textwidth]{interval_avg_defender_payoff_filt_topologyRandom_Erdos_Renyinum_nodes10connectivity_param0p4_matchup_matrix_heatmap.png}
  \caption{Interval-averaged defender payoff across strategy matchups.}
  \label{fig:def_payoff_heatmap}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.45\textwidth]{interval_detection_rate_filt_topologyRandom_Erdos_Renyinum_nodes10connectivity_param0p4_matchup_matrix_heatmap.png}
  \caption{Detection rate across strategy matchups.}
  \label{fig:detection_heatmap}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.45\textwidth]{current_network_health_filt_topologyRandom_Erdos_Renyinum_nodes10connectivity_param0p4_matchup_matrix_heatmap.png}
  \caption{Network health across strategy matchups.}
  \label{fig:health_heatmap}
\end{figure}
%--- End Figures ---

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.45\textwidth]{conv_QL_vs_Bayes_def_defender_atk_QLearning_def_BayesianGame_topo_Random_Erdos_Reny_N10_C0p4.png}
  \caption{Defender payoff convergence: Q-Learning vs Bayesian Game.}
  \label{fig:conv_db}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.45\textwidth]{interval_avg_defender_payoff_attacker_model_huedefender_model_filt_topologyRandom_Erdos_Renyinum_nodes10connectivity_param0p4_boxplot.png}
  \caption{Distribution of final defender payoffs across strategies.}
  \label{fig:def_box}
\end{figure}

@@

\begin{table}[htbp]
  \centering
  \begin{tabular}{lccc}
    \toprule
    \textbf{Metric} & \textbf{F-Statistic} & \textbf{p-value} & \textbf{Significance} \\
    \midrule
    Attacker Payoff & 249.354 & $<$0.001 & Yes \\
    Defender Payoff & 25.065 & $<$0.001 & Yes \\
    Network Health   & 6.375  & $<$0.001 & Yes \\
    Detection Rate   & 90.339 & $<$0.001 & Yes \\
    \bottomrule
  \end{tabular}
  \caption{ANOVA results across strategy matchups for key metrics.}
  \label{tab:anova}
\end{table}

@@

%======================== Code Availability ========================
\section*{Code Availability}
The full code, including \texttt{core\_sim.py}, \texttt{run\_experiments.py}, and \texttt{plot\_results.py}, is available at:\\
\url{https://github.com/umar-projects/jamming-game-theory}. Detailed instructions for replication and dataset generation are provided.

@@

>>>>

% LaTeX template for "A Comparative Study of Game-Theoretic and Reinforcement Learning Approaches for Jam-Resilient Wireless Networks in IoT"
\documentclass[conference]{IEEEtran}
%======================== Packages ========================
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{algorithm,algpseudocode}
\usepackage{cite}
\usepackage{url}
\usepackage{booktabs} % For better tables

\graphicspath{{figures/}} % Tells LaTeX to look in the 'figures' subdirectory

@@

\author{
  \IEEEauthorblockN{Muhammad Umar Yaksambi}
  \IEEEauthorblockA{
    Dept. of CSE (Data Sciences)\\
    RV College Of Engineering\\
    Bengaluru, India\\
    Email: umaryaksambi@gmail.com
  }
  \and
  \IEEEauthorblockN{Anuj Devpura}
  \IEEEauthorblockA{
    Dept. of CSE (Data Sciences)\\
    RV College Of Engineering\\
    Bengaluru, India\\
    Email: anujdevpura2004@gmail.com
  }
  \and
  \IEEEauthorblockN{Sadashiv Todakar}
  \IEEEauthorblockA{
    Dept. of CSE (Data Sciences)\\
    RV College Of Engineering\\
    Bengaluru, India\\
    Email: sadashivtodakar@gmail.com
  }
}

@@

%--- Figures ---
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.45\textwidth]{fig_atk_payoff_heatmap.pdf}
  \caption{Interval-averaged attacker payoff across strategy matchups.}
  \label{fig:atk_payoff_heatmap}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.45\textwidth]{fig_def_payoff_heatmap.pdf}
  \caption{Interval-averaged defender payoff across strategy matchups.}
  \label{fig:def_payoff_heatmap}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.45\textwidth]{fig_detection_heatmap.pdf}
  \caption{Detection rate across strategy matchups.}
  \label{fig:detection_heatmap}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.45\textwidth]{fig_net_health_heatmap.pdf}
  \caption{Network health across strategy matchups.}
  \label{fig:health_heatmap}
\end{figure}
%--- End Figures ---

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.45\textwidth]{fig_def_convergence.pdf}
  \caption{Defender payoff convergence: Q-Learning vs Bayesian Game.}
  \label{fig:conv_db}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.45\textwidth]{fig_def_payoff_boxplot.pdf}
  \caption{Distribution of final defender payoffs across strategies.}
  \label{fig:def_box}   
\end{figure}

@@

\begin{table}[htbp]
  \centering
  \caption{ANOVA results across strategy matchups for key metrics.}
  \label{tab:anova}
  \begin{tabular}{lccc}
    \toprule
    \textbf{Metric} & \textbf{F-Statistic} & \textbf{p-value} & \textbf{Significance} \\
    \midrule
    Attacker Payoff & 249.354 & $<$0.001 & Yes \\
    Defender Payoff & 25.065 & $<$0.001 & Yes \\
    Network Health   & 6.375  & $<$0.001 & Yes \\
    Detection Rate   & 90.339 & $<$0.001 & Yes \\
    \bottomrule
  \end{tabular}
\end{table}

@@

%======================== Code Availability ========================
\section*{Code Availability}
The full code, including \texttt{core\_sim.py}, \texttt{run\_experiments.py}, and \texttt{plot\_results.py}, is available at:\\
\url{https://github.com/UmarYaksambi/Adaptive-Defense-Wireless-Networks}. Detailed instructions for replication and dataset generation are provided.

@@

%======================== References ========================
% Add your bibliography (references.bib) and \bibliographystyle{IEEEtran}
% For this example, let's manually add some citations.
\begin{thebibliography}{1}
\bibitem{axell2012detection} E. Axell, G. Leus, E. G. Larsson, and H. V. Poor, ``Detection and identification of primary users in cognitive radio networks,'' \emph{IEEE Signal Processing Magazine}, vol. 29, no. 3, pp. 113--116, 2012.
\bibitem{chen2019anti} X. Chen and L. Tong, ``Anti-jamming frequency hopping with energy constraint in cognitive radio networks,'' \emph{IEEE Transactions on Vehicular Technology}, vol. 68, no. 1, pp. 317--327, 2019.
\bibitem{alpcan2007game} T. Alpcan and T. Ba\c{s}ar, ``A game theoretic approach to decision and analysis in network intrusion detection,'' in \emph{Proc. 46th IEEE Conf. Decision Control}, 2007, pp. 2595--2602.
\bibitem{khouzani2012evolving} M. H. R. Khouzani and S. Sarkar, ``Evolving games between an attacker and a defender for network security,'' \emph{IEEE/ACM Transactions on Networking}, vol. 20, no. 1, pp. 16--29, 2012.
\bibitem{peng2017anti} C. Peng, H. Li, J. Wang, and K. J. R. Liu, ``Anti-jamming communication based on Q-learning in cognitive radio networks,'' in \emph{Proc. IEEE Int. Conf. Communications (ICC)}, 2017, pp. 1--6.
\bibitem{luo2018anti} Y. Luo, G. Y. Li, W. K. Ma, and H. V. Poor, ``Anti-jamming deep reinforcement learning for spectrum sharing in cognitive radio networks,'' \emph{IEEE Transactions on Cognitive Communications and Networking}, vol. 4, no. 4, pp. 813--824, 2018.
\bibitem{zhao2020hybrid} N. Zhao, F. R. Yu, H. Sun, and M. Li, ``Hybrid game-theoretic and deep reinforcement learning-based anti-jamming for vehicular networks,'' \emph{IEEE Transactions on Vehicular Technology}, vol. 69, no. 8, pp. 8991--9002, 2020.
\bibitem{li2021game} X. Li, M. Zhao, M. Fitch, M. K. K. H. L. L. Chow, and S. Zhang, ``Game theory meets reinforcement learning: A survey of applications in communication networks,'' \emph{IEEE Access}, vol. 9, pp. 67022--67054, 2021.
\end{thebibliography}
